Documentation FFI Sonar Simulator

### 2013-05-23 FFI-Simulator/ffi-simulator.tex
la2pdf commit - 2013:05:23	19:22:24	(Thu23.May)

To simulate this scene we combined a normal optical simulation with a mask image. The mask image scene was set up with a powerful light source and highly diffractive surfaces, which caused everything that was hit by light to obtain a high intensity value. Only direct light was processed, so shadows came out as empty. In the simulated object scene materials were adjusted to reflect a more reasonable amount of light, and a light insensitive background was chosen to avoid double shadows. The two images were finally combined using RGB multiplication.

We believe these simulations to be similar to be similar to the ones initially supplied by FFI, made in Matlab. But are these models accurate enough? Or do we need to consider layover, multipath, or other effects as well?

@@

If layover and multipath are important, then we believe the techniques used to simulate the previous images can not be used. A CUDA / OpenCL implementation can handle this, but the proper way to design it depends on the effects we wish to implement. For layover one could launch a thread for each pixel in the image, which may proceed by finding all patches in the image that contributes by direct returns to that pixel (fig \ref{single_path}). For multipath a ray tracer customized for sound is likely the better option.

\begin{figure}[b]
\graphicsAI[drawing,width=\linewidth]{gfx/layover.svg}
\caption{A CUDA / OpenCL implementation able to handle single path return.}\label{single_path}
\end{figure}

The main advantage of the CUDA implementation lies in its flexibility. If desired, it can be modified to handle sound propagation in non-ideal mediums, penetrable objects, signal multipath, diffraction and phase-dependent effects such as speckle. The drawback is the added code complexity and development time.

We have compared the two methods in Table \ref{comparison}.

>>>>

To simulate this scene we combined a normal optical simulation with a mask image. The mask image scene was set up with a powerful light source and highly diffractive surfaces, which caused everything that was hit by light to obtain a high intensity value. Only direct light was processed, so shadows came out as empty. In the simulated object scene materials were adjusted to reflect a more reasonable amount of light, and a light insensitive background with some added noise was chosen to avoid double shadows. The two images were finally combined using RGB multiplication.

We believe these simulations to be similar to the ones initially supplied by FFI, which were made in Matlab. But are these models accurate enough? Or do we need to consider layover, multipath, or other effects as well?

@@

If layover and multipath are important, then we believe the techniques used to simulate the previous images can not be used. In this case we are likely better off using CUDA / OpenCL from the beginning. For layover one could launch a thread for each pixel in the image, which may proceed by finding all patches in the image that contributes by direct returns to that pixel (fig \ref{single_path}). For multipath a ray tracer altered to handle some sound effects is likely the better option.

\begin{figure}[b]
\graphicsAI[drawing,width=\linewidth]{gfx/layover.svg}
\caption{A CUDA / OpenCL implementation able to handle direct returns and arbitrary geometries. Might approximate the effects of multipath and layover fairly well.}\label{single_path}
\end{figure}

We believe the added flexibility of a CUDA / OpenCL based approach will also make it easier to add new features to the implementation at a later stage, should this be deemed necessary. It could e.g. be altered to handle sound propagation in non-ideal mediums, penetrable objects, signal multipath, diffraction and phase-dependent effects such as speckle. The drawback is the added code complexity and development time.

We also suggest using OpenCL these days, since it is a standard promoted by major players such as Intel and AMD, and because it allows the code to be almost as portable as CUDA. In terms of syntax and maturity the two standards are fairly similar.

We have compared the two alternatives in Table \ref{comparison}.
